---
title: "Chapter 2"
author: "Dominik Klepl"
date: "8 2 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup}
library(rethinking)
```

## R code 2.1
Bag of 4 marbles (blue and white) and sequence BWB was drawn - number of ways the sequence could be drawn assuming all proportions of blue and white marbles:

4xW then there is 0 ways - 0x4x0
1xB+3xW => 1x3x1 = 3 ways
2xW => 2x2x2 = 8
3xB+1xW => 3x1x3 = 9
4xB => 4x0x4 = 0

```{r}
ways <- c( 0 , 3 , 8 , 9 , 0 ) #possible ways each proportion can produce given sequence
ways/sum(ways) #how plausible is each proportion of marbles? also probabilities because they are standardized to add up to 1 together
```

Paramater = proportion of marbles p

Likelihood = number of ways p can produce the data (enumerating all possible data sequences and eliminating the inconsistent ones)

Prior probability = prior plausibility of any specific p

Posterior probability = new, updated plausibility of any specific p given the data

## R code 2.2 (p. 33)
Tossing a globe and counting landings on water and land - 50-50 chance 
How big is the likelihood of getting 6x water in sample of 9
```{r}
dbinom( 6, size=9 , prob=0.5 ) #number of W observations, number of tosses, probability of water in each toss (tosses are independent)

#why is it the same???
A=dbinom( 5, size=9 , prob=0.5 )
B=dbinom( 4, size=9 , prob=0.5 )
A==B
```

## R code 2.3 (p. 40) - Grid approximation
Estimating posterior distribution of the globe tossing example

With prior as a straight line = 1
```{r}
# define grid
p_grid = seq( from=0 , to=1 , length.out=100) #sequence of probability of water from 0 (no water) to 1 (only water) using 20 points on the grid
# 

# define prior
prior = rep (1 , length(p_grid) ) #just a straight line => 20x1

# compute likelihood at each value in grid
likelihood = dbinom( 6 , size=9 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior = likelihood * prior #with prior =1 the posterior=likelihood

# standardize the posterior, so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

#now plot the posterior distribution
plot( p_grid , posterior , type="b" ,
      xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )

#redo this with ggplot
library(ggplot2)
#put all data together into one df
line_posterior = as.data.frame(cbind(p_grid,posterior))
ggplot(line_posterior,aes(p_grid,posterior))+
  geom_line()+
  ggtitle(label="20 points")+
  xlab("probability of water")+
  ylab("posterior probability") 
```

Let's try some different priors

First step prior function
```{r}
prior = ifelse( p_grid < 0.5 , 0 , 1 )

# compute product of likelihood and prior
unstd.posterior = likelihood * prior #with prior =1 the posterior=likelihood

# standardize the posterior, so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

step_posterior = as.data.frame(cbind(p_grid,posterior))

ggplot(step_posterior,aes(p_grid,posterior))+
  geom_line()+
  ggtitle(label="20 points")+
  xlab("probability of water")+
  ylab("posterior probability") 
```

And spike
```{r}
prior = exp( -5*abs( p_grid - 0.5 ) )

# compute product of likelihood and prior
unstd.posterior = likelihood * prior #with prior =1 the posterior=likelihood

# standardize the posterior, so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

spike_posterior = as.data.frame(cbind(p_grid,posterior))

ggplot(spike_posterior,aes(p_grid,posterior))+
  geom_line()+
  ggtitle(label="20 points")+
  xlab("probability of water")+
  ylab("posterior probability") 
```



## R code 2.6 - Quadratic Approximation
1) find the posterior mode (peak of the distribution)
2) estimate curvature near the peak - from that whole posterior distribution is estimated


```{r}
#map = maximum a posteriori
globe.qa = map(
    alist(
        w ~ dbinom(9,p) ,  # binomial likelihood
        p ~ dunif(0,1)     # uniform prior
    ) ,
    data=list(w=6) )

# display summary of quadratic approximation
precis( globe.qa )
```
mean = peak of the distribution = maximum posterior
SD = standard deviation of the posterior distribution
+89% confidence interval

Assuming the posterior is Gaussian, it is maximized at 0.67, and its SD is 0.16

Analytical calculation (look up explanation somewhere else, maybe lecture???) that gives an exact right answer to compare with performance of quadratic approximation
```{r}
## R code 2.7
# analytical calculation
w = 6
n = 9
curve( dbeta( x , w+1 , n-w+1 ) , from=0 , to=1 )
# quadratic approximation
curve( dnorm( x , 0.67 , 0.16 ) , lty=2 , add=TRUE )
```

#PRACTICE PROBLEMS (p. 45-47)

##MEDIUM 1
```{r}
#build function to make this faster
get_posterior = function (obs, size,prior,grid) {
  likelihood = dbinom( obs , size=size , prob=grid )
  # compute product of likelihood and prior
  unstd.posterior = likelihood * prior #with prior =1 the posterior=likelihood
  # standardize the posterior, so it sums to 1
  posterior = unstd.posterior / sum(unstd.posterior)
  df_posterior = as.data.frame(cbind(grid,posterior))
  
  print(ggplot(df_posterior,aes(grid,posterior))+
          geom_line()+
          ggtitle(label="100 points")+
          xlab("probability of water")+
          ylab("posterior probability")
  )

return (df_posterior)
}

#prior for this problem is the same for all
prior = rep (1 , length(p_grid) ) #just a straight line => 20x1
#and grid
p_grid = seq( from=0 , to=1 , length.out=100)

#1. WWW
answer=get_posterior(3,3,prior,p_grid)
answer$grid[answer$posterior==max(answer$posterior)]

#2. WWWL
answer=get_posterior(obs=3,size=4,prior,p_grid)
answer$grid[answer$posterior==max(answer$posterior)]


#3. LWWLWWW
answer=get_posterior(obs=5,size=7,prior,p_grid)
answer$grid[answer$posterior==max(answer$posterior)]
```

##MEDIUM 2
```{r}
#now as in 1 just with step prior
prior = ifelse(test = p_grid < .5, yes = 0, no = 1)

#1. WWW
answer=get_posterior(3,3,prior,p_grid)
answer$grid[answer$posterior==max(answer$posterior)]

#2. WWWL
answer=get_posterior(obs=3,size=4,prior,p_grid)
answer$grid[answer$posterior==max(answer$posterior)]


#3. LWWLWWW
answer=get_posterior(obs=5,size=7,prior,p_grid)
answer$grid[answer$posterior==max(answer$posterior)]
```


##MEDIUM 3
```{r}
prior = c(.5, .5)
likelihood = c(.3, 1)
unstd_posterior = prior*likelihood
posterior = unstd_posterior / sum(unstd_posterior)
posterior
```

##MEDIUM 4
```{r}
#n of ways black side could be produced
  #2 black sides => 2
  #1 black 1 white => 1
  #2 whites => 0

likelihood = c(2, 1, 0)
prior = rep (1, 3)
unstd_posterior = likelihood*prior

posterior = unstd_posterior/sum(unstd_posterior)

posterior[1] ==2/3
```

##MEDIUM 5
```{r}
#now I have two black-black cards so posterior change for them as they are twice as likely to be drawn
prior[1] = prior[1]*2
unstd_posterior = likelihood*prior

posterior = unstd_posterior/sum(unstd_posterior)

posterior[1]
```

##MEDIUM 6
```{r}
#likelihood still the same
#prior changes
prior = c(1,2,3)

unstd_posterior = likelihood*prior

posterior = unstd_posterior/sum(unstd_posterior)

posterior[1] ==0.5
```

##MEDIUM 7
```{r}
card.1.2.likelihood = 2
card.2.1.likelihood = 0
card.1.3.likelihood = 4
card.3.1.likelihood = 0
card.2.3.likelihood = 2
card.3.2.likelihood = 0

likelihood = c(card.1.2.likelihood, card.2.1.likelihood, card.1.3.likelihood, card.3.1.likelihood, card.2.3.likelihood, card.3.2.likelihood)
prior = rep(x = 1, length = length(likelihood))
unstandardized.posterior = prior * likelihood
posterior = unstandardized.posterior / sum(unstandardized.posterior)

# the probability that the other side of the first card is black is equal to the probability that the first card is card 1,
# which equals the probability that the sequence we've chosen is either (1, 2), or (1, 3)
posterior[1] + posterior[3] == 0.75
```

##HARD 1
```{r}
species_A = 0.1
species_B = 0.2

likelihood = c(species_A,species_B)

prior = c(0.5,0.5)
unstandardized.posterior = prior * likelihood
posterior = unstandardized.posterior / sum(unstandardized.posterior)

posterior[1]*.1 + posterior[2]*.2
```

##HARD 2
```{r}
posterior[1]
```

##HARD 3
```{r}
# First birth = Twins
# Second birth = Single
# P(Species = A| Twins, Single)?

# P(Species = A | Twins, Single) = P(Twins | Species = A) *
#                                   P(Single | Species = A) *
#                                      P(A) /
#                                         P(Twins,Single)
# P(Twins,Single) =  P(Twins | Species = A) *
#                       P(Single | Species = A) *
#                          P(A) +
#                    P(Twins | Species = B) *
#                       P(Single | Species = B) *
#                          P(B)

p_A_twins_single = (0.1 * 0.9 * 0.5) / ((0.1 * 0.9 * 0.5) + (0.2 * 0.8 * 0.5))

p_A_twins_single
```

##HARD 4

###HARD 4.1
```{r}
#first just the test
prior = c(0.5,0.5)

likelihood = c(.8,1-.65)

unstandardized.posterior = prior * likelihood
posterior = unstandardized.posterior / sum(unstandardized.posterior)

posterior[1]
```
###HARD 4.2
```{r}
#posterior from 4.1 becomes prior now
prior = posterior

#and add the likelihood from H3 - probability of both species having twins and then single
species_A = 0.1 * (1-0.1)
species_B = 0.2 * (1-0.2)

likelihood = c(species_A,species_B)

unstandardized.posterior = prior * likelihood
posterior = unstandardized.posterior / sum(unstandardized.posterior)

posterior[1]
```

#BONUS
Show the change of in probability with adding evidence

```{r}
evidence = factor(rep(c("prior", 
                        "prior + twins", 
                        "prior + twins + single", 
                        "prior + twins + single + test = A"), 
               each = 2), 
               levels = c("prior", 
                        "prior + twins", 
                        "prior + twins + single", 
                        "prior + twins + single + test = A"))
species = rep(c("Species A", "Species B"), 4)
prob = c(0.5, 0.5, 0.33, 0.67, 0.36, 0.64, 0.57, 0.43)
change_df <- data.frame(species = species,
                        evidence = evidence,
                        prob = prob)
ggplot(change_df, aes(x = evidence, 
                      y = prob, 
                      colour = species, 
                      group = species)) +
  geom_line() +
  geom_point(size = 3) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab("Data") +
  ylab("Posterior Probability")
```


