---
title: "Chapter 2"
author: "Dominik Klepl"
date: "8 2 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup}
library(rethinking)
```

## R code 2.1
Bag of 4 marbles (blue and white) and sequence BWB was drawn - number of ways the sequence could be drawn assuming all proportions of blue and white marbles:

4xW then there is 0 ways - 0x4x0
1xB+3xW => 1x3x1 = 3 ways
2xW => 2x2x2 = 8
3xB+1xW => 3x1x3 = 9
4xB => 4x0x4 = 0

```{r}
ways <- c( 0 , 3 , 8 , 9 , 0 ) #possible ways each proportion can produce given sequence
ways/sum(ways) #how plausible is each proportion of marbles? also probabilities because they are standardized to add up to 1 together
```

Paramater = proportion of marbles p

Likelihood = number of ways p can produce the data (enumerating all possible data sequences and eliminating the inconsistent ones)

Prior probability = prior plausibility of any specific p

Posterior probability = new, updated plausibility of any specific p given the data

## R code 2.2 (p. 33)
Tossing a globe and counting landings on water and land - 50-50 chance 
How big is the likelihood of getting 6x water in sample of 9
```{r}
dbinom( 6, size=9 , prob=0.5 ) #number of W observations, number of tosses, probability of water in each toss (tosses are independent)

#why is it the same???
A=dbinom( 5, size=9 , prob=0.5 )
B=dbinom( 4, size=9 , prob=0.5 )
A==B
```

## R code 2.3 (p. 40) - Grid approximation
Estimating posterior distribution of the globe tossing example

With prior as a straight line = 1
```{r}
# define grid
p_grid = seq( from=0 , to=1 , length.out=100) #sequence of probability of water from 0 (no water) to 1 (only water) using 20 points on the grid
# 

# define prior
prior = rep (1 , length(p_grid) ) #just a straight line => 20x1

# compute likelihood at each value in grid
likelihood = dbinom( 6 , size=9 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior = likelihood * prior #with prior =1 the posterior=likelihood

# standardize the posterior, so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

#now plot the posterior distribution
plot( p_grid , posterior , type="b" ,
      xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )

#redo this with ggplot
library(ggplot2)
#put all data together into one df
line_posterior = as.data.frame(cbind(p_grid,posterior))
ggplot(line_posterior,aes(p_grid,posterior))+
  geom_line()+
  ggtitle(label="20 points")+
  xlab("probability of water")+
  ylab("posterior probability") 
```

Let's try some different priors

First step prior function
```{r}
prior = ifelse( p_grid < 0.5 , 0 , 1 )

# compute product of likelihood and prior
unstd.posterior = likelihood * prior #with prior =1 the posterior=likelihood

# standardize the posterior, so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

step_posterior = as.data.frame(cbind(p_grid,posterior))

ggplot(step_posterior,aes(p_grid,posterior))+
  geom_line()+
  ggtitle(label="20 points")+
  xlab("probability of water")+
  ylab("posterior probability") 
```

And spike
```{r}
prior = exp( -5*abs( p_grid - 0.5 ) )

# compute product of likelihood and prior
unstd.posterior = likelihood * prior #with prior =1 the posterior=likelihood

# standardize the posterior, so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

spike_posterior = as.data.frame(cbind(p_grid,posterior))

ggplot(spike_posterior,aes(p_grid,posterior))+
  geom_line()+
  ggtitle(label="20 points")+
  xlab("probability of water")+
  ylab("posterior probability") 
```



## R code 2.6 - Quadratic Approximation
1) find the posterior mode (peak of the distribution)
2) estimate curvature near the peak - from that whole posterior distribution is estimated


```{r}
#map = maximum a posteriori
globe.qa = map(
    alist(
        w ~ dbinom(9,p) ,  # binomial likelihood
        p ~ dunif(0,1)     # uniform prior
    ) ,
    data=list(w=6) )

# display summary of quadratic approximation
precis( globe.qa )
```
mean = peak of the distribution = maximum posterior
SD = standard deviation of the posterior distribution
+89% confidence interval

Assuming the posterior is Gaussian, it is maximized at 0.67, and its SD is 0.16

Analytical calculation (look up explanation somewhere else, maybe lecture???) that gives an exact right answer to compare with performance of quadratic approximation
```{r}
## R code 2.7
# analytical calculation
w = 6
n = 9
curve( dbeta( x , w+1 , n-w+1 ) , from=0 , to=1 )
# quadratic approximation
curve( dnorm( x , 0.67 , 0.16 ) , lty=2 , add=TRUE )
```